News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-24 rupnik2015.tex
Small fixes of typos.

The system is designed to handle streams and so we describe how a document which enters the stream is processed. The sequence of operations is shown in  Figure~\ref{fig:erpipeline}.  A new document is first tokenized, stop words are removed and words are stemmed.
 What remains is represented in a vector-space model and normalized using $TF-IDF$ (see Section~\ref{sec:tfidf} for the definition). Since each article is tagged with its language, cosine similarity  is used to find the most similar existing cluster, by comparing it to the centroid vector of each cluster. A selected threshold is used to determine if the article is not similar to any existing clusters (0.4 was used in our experiments).  The article is then assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

The main application on which we test the above similarity is cross-lingual event linking. In online media streams -- particularly new articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. Same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to find suitable matchings between clusters with articles in different languages.

@@

To investigate the empirical performance of the low rank approximations we will test the algorithms on a large-scale, real-world multi-lingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multi-lingual data that is especially important for the languages for which no translation tools, multi-lingual dictionaries as Eurovoc, or strongly aligned multi-lingual corpora as Europarl are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million%$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10 k articles, and $236$ with more than $1000$ articles.

@@

 In this subsection, we will investigate the empirical performance of CCA hub approach. We will demonstrate that this approach can be successfully applied even in the case of fully missing alignment information.
 To this purpose, we select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing, we remove the documents that contain less than 20 different words (stubs) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors.
 %These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English %language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too %infrequent as well as the top 100 most frequent words in the vocabularies.
 Although the English language is well aligned with all Wikipedia languages, we must note that quality of alignment varies quite a bit, especially in the case of small Wikipedias.
 Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.

@@

\subsection{Evaluation Of Cross-Lingual Event Linking}
In order to determine how accurately we can predict cluster equivalence, we performed two experiments in multi-lingual setting using English, German and Spanish language for which we had labelled data to evaluate the linking performance. In the first experiment, we tested how well  the individual approaches for cross-lingual article linking perform when used for linking the clusters about the same event. In the second experiment we tested how accurate is the prediction model when trained on different subsets of learning features. To evaluate the prediction accuracy for a given dataset we used 10-fold cross validation. In this technique, the dataset is partitioned into 10 folds, where 9 folds are used for training a model and the left-out fold is then used to test the accuracy of the model. The learning and testing process is repeated 10 times, each time with a different left-out fold.

@@

% AMMR experiments
We also compared the proposed approaches on the task of Wikipedia mate retrieval (the same task as in Section~\ref{experiments:hubcca}). We computed the Average (over language pairs) Mean Reciprocal Rank (AMRR) \shortcite{voorhees1999trec}  performance of the different approaches on the  Wikipedia data by holding out $15000$ aligned test documents and using $300000$ aligned documents as the training set. The Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that hub CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. The hub CCA approach manages with $500$ features performam comparably to $LSI$ based approach with $1000$-features, which shows that the $CCA$ method can improve both model memory footprint as well as similarity computation time.

@@

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support any language from the top $100$ Wikipedia languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) which provides $430k$ unique articles per day. Our system currently computes 400k similarities per 200ms, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for each language which requires 1.5 GB of memory with documents   stored as 500 dimensional vectors. We  note that the time complexity of the similarity computations scales linearly with dimension of the feature space and does not  depend of number of languages. For each article, we compute the top $10$  most similar ones in every other language.

@@

\item \textbf{Features} - The experiments show that the hub CCA based features present good baseline, which can greatly benefit from additional semantic extraction based features. Even though the CCA based features do not greatly improve the system performance when the semantic features are used, there are two main benefits in the approach: the linking process can be sped up (smaller candidate cluster sets), and in the case when semantic extraction is not possible.

>>>>

The system is designed to handle streams and so we describe how a document which enters the stream is processed. The sequence of operations is shown in  Figure~\ref{fig:erpipeline}.  A new document is first tokenized, stop words are removed and words are stemmed.
 What remains is represented in a vector-space model and normalized using $TFIDF$ (see Section~\ref{sec:tfidf} for the definition). Since each article is tagged with its language, cosine similarity  is used to find the most similar existing cluster, by comparing it to the centroid vector of each cluster. A selected threshold is used to determine if the article is not similar to any existing clusters (0.4 was used in our experiments).  The article is then assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

The main application on which we test the above similarity is cross-lingual event linking. In online media streams -- particularly new articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. Same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to finding suitable matchings between clusters with articles in different languages.

@@

To investigate the empirical performance of the low rank approximations we will test the algorithms on a large-scale, real-world multi-lingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multi-lingual data that is especially important for the languages for which no translation tools, multi-lingual dictionaries as Eurovoc, or strongly aligned multi-lingual corpora as Europarl are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1000$ articles.

@@

 In this subsection, we will investigate the empirical performance of hub CCA approach. We will demonstrate that this approach can be successfully applied even in the case of fully missing alignment information.
 To this purpose, we select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing, we remove the documents that contain less than 20 different words (stubs) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors.
 %These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English %language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too %infrequent as well as the top 100 most frequent words in the vocabularies.
 Although the English language is well aligned with all Wikipedia languages, we must note that quality of alignment varies quite a bit, especially in the case of small Wikipedias.
 Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.

@@

\subsection{Evaluation Of Cross-Lingual Event Linking}
In order to determine how accurately we can predict cluster equivalence, we performed two experiments in multi-lingual setting, English, German and Spanish language for which we had labelled data to evaluate the linking performance. In the first experiment, we tested how well  the individual approaches for cross-lingual article linking perform when used for linking the clusters about the same event. In the second experiment we tested how accurate is the prediction model when trained on different subsets of learning features. To evaluate the prediction accuracy for a given dataset we used 10-fold cross validation. In this technique, the dataset is partitioned into 10 folds, where 9 folds are used for training a model and the left-out fold is then used to test the accuracy of the model. The learning and testing process is repeated 10 times, each time with a different left-out fold.

@@

% AMMR experiments
We also compared the proposed approaches on the task of Wikipedia mate retrieval (the same task as in Section~\ref{experiments:hubcca}). We computed the Average (over language pairs) Mean Reciprocal Rank (AMRR) \shortcite{voorhees1999trec}  performance of the different approaches on the  Wikipedia data by holding out $15000$ aligned test documents and using $300000$ aligned documents as the training set. The Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that hub CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. The hub CCA approach with $500$ features manages to perform comparably to $LSI$ based approach with $1000$-features, which shows that the $CCA$ method can improve both model memory footprint as well as similarity computation time.

@@

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support any language from the top $100$ Wikipedia languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) which provides $430k$ unique articles per day. Our system currently computes 400k similarities per 200ms, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for each language which requires 1.5 GB of memory with documents   stored as 500 dimensional vectors. We  note that the time complexity of the similarity computations scales linearly with dimension of the feature space and does not  depend on number of languages. For each article, we compute the top $10$  most similar ones in every other language.

@@

\item \textbf{Features} - The experiments show that the hub CCA based features present good baseline, which can greatly benefit from additional semantic extraction based features. Even though the CCA based features do not greatly improve the system performance when the semantic features are used, there are two main benefits in the approach: the linking process can be sped up (smaller candidate cluster sets), and usage in the case when semantic extraction is not possible.
